Gravitational Lens Modelling of Flux Ratio Anomalous System B1555+375

### 2015-04-23 papers/B1555_EdgeOn/B1555_letter.tex
minor changes

A common feature of simulations of structure formation is the presence
of thousands of subhaloes that are associated with larger mass haloes \citep[e.g.][]{Springel08}.
However, observations of the Local Group find many fewer satellite
galaxies than are predicted by the simulations, even when taking into
account completeness corrections arising from the limited sky coverage
imposed by the Galactic Plane {\bf and the latest discovery of new dwarf galaxies} \citep{DES15,Kop15}.  
This is the famous ``missing satellite'' problem \citep{Klypin1999, Moore1999, S07}. In order to
understand this discrepancy, it is imperative to build up a large
sample of satellites/subhaloes in galaxies outside the Local Group.
{\bf However,} this process is challenging due to the faintness of the satellite
galaxies.  Furthermore, lower-mass satellites may not be able to
retain the gas needed for ongoing star formation \citep[e.g.,][]{P11},
rendering them effectively dark.  These factors make gravitational
lensing a powerful and promising tool for the detection of
substructure.

@@

While future large-scale surveys, such as the LSST, are expected to
deliver new large samples of gravitationally lensed quasars, in the
near future the discovery of new systems using narrow-line quasar
emission \citep{N14} will provide new invaluable statistical
perspective on current constraints based on flux ratio anomalies.  In
this {\bf letter}, we explore the idea that not all flux ratio anomalies are
related to the presence of substructure, but that some originate in a
more complex lens galaxy mass distribution than initially
considered. To this end we use new infra-red data from the Strong
lensing at High Angular Resolution Program (SHARP; Fassnacht et al.,
in prep). The main goal of the SHARP project is to re-observe known
quadruple and Einstein ring lens systems at a much higher angular
resolution by making use of observations with (1) adaptive optics (AO)
on the Keck telescopes, (2) the Hubble Space Telescope, and (3) large
interferometric arrays \citep{SHARP12,V12}. In particular, to better
understand the causes of flux ratio anomalies in gravitational lenses
we focus on the B1555+375 lens system\citep{Marlow99}.  This system was
discovered in the Cosmic Lens All-Sky Survey \citep{CLASS1,CLASS2} and
has a pair of merging images that show a strong flux ratio anomaly at
radio wavelengths.  Follow-up observations ruled out radio propagation
effects as the cause of the anomaly \citep{K03,KD04}.  The B1555+375
system was included in both the observational and theoretical analysis
of \citet{Dalal2002} and \citet{Xu14} and, thus, understanding its
mass distribution is of key importance.  Our analysis makes use of
the new information obtained from the latest SHARP infrared imaging.
The paper is organized as follows: the latest SHARP data are
introduced in section 2; in section 3 we present the lens model while
the results of this analysis and their implications are discussed in
section 4.

@@

Gravitational lensing is, {\bf at present,} the most powerful tool for
studying substructure in galaxies outside of the Local Universe, and
the flux ratio anomaly systems are a key aspect of these
investigations.  However, before using the flux ratio anomaly in any
given system to investigate structure formation and the properties of
dark matter, it is critical to eliminate other possible explanations
for the difference between the observed and expected fluxes{\bf, such as propagation effects and assumptions on the mass distribution of the lens galaxy}.  
It is for this reason that much of the effort in the flux-ratio technique
has been focused on radio-loud lens systems.  At these wavelengths,
dust extinction and microlensing effects should be minimal
\citep[although see][for a rare example of radio microlensing]{K2000}.
Radio propagation effects can affect radio fluxes, as is clearly shown
in the B0128+437 system \citep{B04}.  However, these propagation
effects can be disentangled from substructure lensing because they
have well known dependencies on wavelength. \citet{KD04} examine all
eight radio loud flux anomalous lenses besides B0128+437, namely: MG
0414+0534, B0712+472, B1359+154, B1422+231, B1555+375, B1608+656,
B1933+503, and B2045+265.  They found no strong wavelength dependence
in the flux ratios of all these systems and therefore rule out radio
propagation effects as significant contributors to the observed
anomalies.

@@

From a theoretical perspective, \citet{Xu14} find that substructures
cannot explain the observed radio flux anomalies in all the lenses
that were considered. By adding subhaloes from numerical simulations
on to these systems, only one out of eight radio loud systems
reproduces the level of flux anomalies seen in the data.  The rest of
them are unlikely (only $1 \% - 4 \%$ probability) to have flux
anomalies produced by substructure, including our target B1555+375.
In fact, among those eight radio lenses, B1555+375 is one of the most
anomalous systems, but it is also one of the most challenging systems
to explain by substructures. \citet{Xu14} have concluded that the failure in
predicting flux ratios is most likely due to oversimplified/improper lens
modelling. They also point out that the SIE component and the external
shear in their model are orthogonal, indicating a possible missing
ingredient in the lens model. Their results are consistent with the SHARP
$K^{\prime}$ band observations, in which an edge-on disk crosses the
radio merging double and which fully accounts for the observed flux ratio anomaly.  In general, it is well known (e.g,
\citep{Ka91}) that lens models based only on the locations of lensed
quasar/AGN images have large degeneracies. Therefore, before
concluding that the flux ratios in any given lens system are a result
of substructure, it is critical to explore a range of smooth models
that are informed by additional observations or to do a comprehensive
analysis such as that conducted by \citet{Xu14}. Otherwise, derived
substructure fractions should be considered as upper limits.

Although the number of radio loud flux anomalous lensed
quasars remains small, a new technique using narrow-line emission
regions will enlarge the sample and bring statistical perspective on
substructures in the near future \citep{N14}.

>>>>

A common feature of simulations of structure formation is the presence
of thousands of subhaloes that are associated with larger mass haloes \citep[e.g.][]{Springel08}.
However, observations of the Local Group find many fewer satellite
galaxies than are predicted by the simulations, even when taking into
account completeness corrections arising from the limited sky coverage
imposed by the Galactic Plane and the latest discovery of new dwarf galaxies \citep{DES15,Kop15}.  
This is the famous ``missing satellite'' problem \citep{Klypin1999, Moore1999, S07}. In order to
understand this discrepancy, it is imperative to build up a large
sample of satellites/subhaloes in galaxies outside the Local Group.
 However, this process is challenging due to the faintness of the satellite
galaxies.  Furthermore, lower-mass satellites may not be able to
retain the gas needed for ongoing star formation \citep[e.g.,][]{P11},
rendering them effectively dark.  These factors make gravitational
lensing a powerful and promising tool for the detection of
substructure.

@@

While future large-scale surveys, such as the LSST, are expected to
deliver new large samples of gravitationally lensed quasars, in the
near future the discovery of new systems using narrow-line quasar
emission \citep{N14} will provide new invaluable statistical
perspective on current constraints based on flux ratio anomalies.  In
this letter, we explore the idea that not all flux ratio anomalies are
related to the presence of substructure, but that some originate in a
more complex lens galaxy mass distribution than initially
considered. To this end we use new infra-red data from the Strong
lensing at High Angular Resolution Program (SHARP; Fassnacht et al.,
in prep). The main goal of the SHARP project is to re-observe known
quadruple and Einstein ring lens systems at a much higher angular
resolution by making use of observations with (1) adaptive optics (AO)
on the Keck telescopes, (2) the Hubble Space Telescope, and (3) large
interferometric arrays \citep{SHARP12,V12}. In particular, to better
understand the causes of flux ratio anomalies in gravitational lenses
we focus on the B1555+375 lens system\citep{Marlow99}.  This system was
discovered in the Cosmic Lens All-Sky Survey \citep{CLASS1,CLASS2} and
has a pair of merging images that show a strong flux ratio anomaly at
radio wavelengths.  Follow-up observations ruled out radio propagation
effects as the cause of the anomaly \citep{K03,KD04}.  The B1555+375
system was included in both the observational and theoretical analysis
of \citet{Dalal2002} and \citet{Xu14} and, thus, understanding its
mass distribution is of key importance.  Our analysis makes use of
the new information obtained from the latest SHARP infrared imaging.
The paper is organized as follows: the latest SHARP data are
introduced in section 2; in section 3 we present the lens model while
the results of this analysis and their implications are discussed in
section 4.

@@

Gravitational lensing is,  at present, the most powerful tool for
studying substructure in galaxies outside of the Local Universe, and
the flux ratio anomaly systems are a key aspect of these
investigations.  However, before using the flux ratio anomaly in any
given system to investigate structure formation and the properties of
dark matter, it is critical to eliminate other possible explanations
for the difference between the observed and expected fluxes, such as propagation effects and assumptions on the mass distribution of the lens galaxy.  
It is for this reason that much of the effort in the flux-ratio technique
has been focused on radio-loud lens systems.  At these wavelengths,
dust extinction and microlensing effects should be minimal
\citep[although see][for a rare example of radio microlensing]{K2000}.
Radio propagation effects can affect radio fluxes, as is clearly shown
in the B0128+437 system \citep{B04}.  However, these propagation
effects can be disentangled from substructure lensing because they
have well known dependencies on wavelength. \citet{KD04} examine all
eight radio loud flux anomalous lenses besides B0128+437, namely: MG
0414+0534, B0712+472, B1359+154, B1422+231, B1555+375, B1608+656,
B1933+503, and B2045+265.  They found no strong wavelength dependence
in the flux ratios of all these systems and therefore rule out radio
propagation effects as significant contributors to the observed
anomalies.

@@

From a theoretical perspective, \citet{Xu14} find that substructures
cannot explain the observed radio flux anomalies in all the lenses
that were considered. By adding subhaloes from numerical simulations
on to these systems, only one out of eight radio loud systems
reproduces the level of flux anomalies seen in the data.  The rest of
them are unlikely (only $1 \% - 4 \%$ probability) to have flux
anomalies produced by substructure, including our target B1555+375.
In fact, among those eight radio lenses, B1555+375 is one of the most
anomalous systems, but it is also one of the most challenging systems
to explain by substructures. \citet{Xu14} have concluded that the failure in
predicting flux ratios is most likely due to oversimplified/improper lens
modelling. They also point out that the SIE component and the external
shear in their model are orthogonal, indicating a possible missing
ingredient in the lens model. Their results are consistent with the SHARP
$K^{\prime}$ band observations, in which an edge-on disk crosses the
radio merging double and which fully accounts for the observed flux ratio anomaly.  In general, it is well known 
\citep[e.g,][]{Ka91} that lens models based only on the locations of lensed
quasar/AGN images have large degeneracies. Therefore, before
concluding that the flux ratios in any given lens system are a result
of substructure, it is critical to explore a range of smooth models
that are informed by additional observations or to do a comprehensive
analysis such as that conducted by \citet{Xu14}. Otherwise, derived
substructure fractions should be considered as upper limits.

%Although the number of radio loud flux anomalous lensed
%quasars remains small, a new technique using narrow-line emission
%regions will enlarge the sample and bring statistical perspective on
%substructures in the near future \citep{N14}.
